---
title: "Orientation / Instructor Notes"
---

## Projects, Notebooks, import/export, display

1. Ask students to create a blank new RStudio Project: "hello world"
1. Create a Blank R Notebook (prose [w R Markdown] & code chunks)
1. Insert Code Chunk and `library(tidyverse) ; library(fs)`
1. fs::dir_create("data")  -- click New Folder ; data
1. Insert Code Chunk and `write_csv(mpg, "data/mpg2.csv")`
1. Insert Code Chunk and `mpg_2 <- read_csv("data/mpg2.csv")`
1. Insert Code Chunk and `mpg_2`
1. Save the file:  hello_notebook.Rmd
1. Preview the file:  *Ctrl+Shift+K* 
1. Click the "Open in Browser" button


```{r}
# In the Console Tab

cars
View(cars)

# See help in the Help tab, type the following in the Console tab
?cars
?View
# See Also, from the Menubar:  Help > Cheatsheets

5 + 5
answer <- 5 + 5
numbers <- seq(3,45, by = 4)
answer
numbers
answer * numbers
answer2 <- (answer / 4) * numbers
answer2
first_names <- c("Neha", "Jinyuan", "Stella")
cars$speed + 5

```


```{r}
library(tidyverse)

mpg %>% 
  ggplot(aes(displ, hwy, color = class)) +
  geom_point()
```

