---
title: "Orientation / Instructor Notes"
---

## Projects, Notebooks, import/export, display

0. Download intro2r-code repo and open the .Rproj
1. Ask students to create a blank new RStudio Project: "hello world"
1. Create a Blank R Notebook (prose [w R Markdown] & code chunks)
1. Save the file:  hello_notebook.Rmd
1. Preview the file:  *Ctrl+Shift+K*
1. switch back to first project: "intro2r-code" 
2. Instructor demo a new notebook


2. Insert Code Chunk and `library(tidyverse) ; library(fs)`
2. fs::dir_create("data")  -- click New Folder ; data

2. Insert Code Chunk and `write_csv(mpg, "data/mpg2.csv")`
2. Insert Code Chunk and `mpg_2 <- read_csv("data/mpg2.csv")`
2. Insert Code Chunk and `mpg_2`

3. Save the file:  hello_notebook.Rmd
3. Preview the file:  *Ctrl+Shift+K* 
3. Click the "Open in Browser" button

> Students should open the intro2r-code project
> Do exercise_00.Rmd (preview and open in browser)
> Answers can be found in answers_00.Rmd or answers_00.nb.html

## RStudio environment and R 

INSTRUCTOR Demo


```{r}
# In the Console Tab

cars
View(cars)
```


```{r}
# See help in the Help tab, type the following in the Console tab
?cars
?View

```


```{r}
5 + 5
```


```{r}
answer <- 5 + 5
```


```{r}
answer 
```


```{r}
numbers <- seq(3,45, by = 4)
numbers
```


```{r}
answer * numbers
```


```{r}
answer2 <- (answer / 4) * numbers
answer2
```


```{r}
first_names <- c("Neha", "Jinyuan", "Stella")
first_names
```


```{r}
paste("Hello", first_names, sep = " ")
```


```{r}
cars
cars$speed + 5
```

## Library Packages

Review

- install once
- load each time
- overview the tidyverse loaded packages

```{r}
library(tidyverse)  #this should be at the top
```

## Bonus visualization

ggplot2 is a visualization pacakge and part of the tidyverse.

```{r}
mpg %>% 
  ggplot(aes(displ, hwy, color = class)) +
  geom_point()
```

## Next

Data Wrangling with 01_dplyr.Rmd.  Use select slides/index.html to illustrate concepts.

Follow with exercise_01.Rmd and answers_01.nb.html

> Optionally, use [00 Literate Programming notebook](00_import-data_literate-programming.nb.html) to reiterate import, export, packages, rendered R Markdown, reports, etc.
