---
title: "my first notebook"
author: "John Little"
date: "April 1, 2020"
output:  html_notebook
abstract: This is an [R Notebook](https://bookdown.org/yihui/rmarkdown/notebook.html) techniques with code **inside** *code chunks*.  Outside the code chunks you can write your prose to explain your goals, workflow, provenance, and code.  You can illustrating [literate programming](https://en.wikipedia.org/wiki/Literate_programming) enhance your prose with [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html)).
---

<p>

You can execute code within the R Notebook. Results appear beneath the code.  R Markdown documents can be rendered into many formats including MS Word, Power Point, slides, web pages, web sites, and more. 

Try executing the **code chunk**, below, by clicking the green-arrow *Run* button (within the chunk), or by placing your cursor inside the chunk and pressing *Ctrl+Shift+Enter.*


```{r}
plot(cars)
```

## Code Chunks

Add a new code chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

*Ctrl+Shift+M*:  Generates a pipe -- ` %>% ` -- between functions (When you use the `tidyerse` package.)

###### https://rmarkdown.rstudio.com/r_notebooks.html

## Load Library Packages 

We like the [tidyverse](https://tidyverse.org).  The tidyverse is a collection of R packages that enable a modern approach to data science and data manipulation.  The packages share a common philosophy and employ a consistent approach to documentation.  Package [documentation can be explored on the web](https://www.tidyverse.org/packages/).

The Tidyverse is a mega package that loads eight packages at once.


```{r}
library(tidyverse)
```

Packages are a means to extend R's functionality into your domain of interest.  The tidyverse is a general interest set of packages concerned with data science and data manipulation.  It just makes R programming easier.

Packages ....

1. Only need to be installed once  `install.packages("tidyverse)`
1. Need to be loaded during every R session   `library(tidyverse)`

Before we go any further, let's export some CSV data so we can practice importing that data later.

```{r}
write_csv(cars, "data/cars2.csv")
```


## R Markdown

Markdown is a minimal "mark up" language that enables document structure affected when you **render** (or *preview*, or *knit*) the R Markdown Notebook script (.Rmd) into your chosen report format (HTML notebook, MS Word file, PDF Document, e-book, web site, dashboard, slides, PPT slides, etc.)

Below is an example of some R Markdown in an R Notebook.  Preview this document to see how it renders.

My Your_Name_Here *favorite* **search** *engines* are 

1. [Google](http://google.com)
3. [Yahoo](http://yahoo.com)
2. [Bing](http://bing.com)

You can get more information about getting started with [R Markdown](https://rmarkdown.rstudio.com/)



## Load Data

The following links provide documentation on loading data into R

- https://support.rstudio.com/hc/en-us/articles/218611977-Importing-Data-with-RStudio
- https://www.rstudio.com/resources/webinars/importing-data-into-r/

### Hurricane data 

Let's load some data.  Below we import hurricane data from the web (from github) and assign that data to the `canes` object.  Next we import a CSV file from the data directory and assign that to the `cars2` 

```{r}
canes <- read_csv("https://is.gd/bigstorms")
cars2 <- read_csv("data/cars2.csv")
```

Now you can see that data by calling 

```{r}
canes
# How do you see the cars2 data frame?
```

have a look at the data structures and data types

###### http://r4ds.had.co.nz/tibbles.html

```{r}
class(canes)
```

###### http://r4ds.had.co.nz/vectors.html

```{r}
glimpse(canes)
```

