---
title: "Answers 01 - dplyr -- data wrangling"
output: html_notebook
---

### Setup

3. Load the `tidyverse` library package

```{r}
library(tidyverse)
```


4.

```{r}
brodhead <- read_csv("data/brodheadCenter.csv")
```

### exercise_01 -- Data Wrangling

1. Which restaurant has the lowest cost item and what is the item?

```{r}
brodhead %>% 
  arrange(cost) %>% 
  select(name, type, itemName, cost)
```

#### Alternative Answer

```{r}
brodhead %>% 
  filter(cost == min(cost)) %>% 
  select(name, type, itemType, cost)
  
```

2. Which restaurant has the most expensive item(s)?  What are those item(s)?

```{r}
brodhead %>% 
  arrange(desc(cost)) %>% 
  select(name, type, itemName, cost)
```

#### Alternative Answer

```{r}
answer2 <- brodhead %>% 
  select(name, type, itemName, cost) %>% 
  filter(type == "restaurant",
         cost == max(cost)) 
answer2
```

#### Example of inline coding (Render report to see result)

The **`r answer2$name[1]`** *`r answer2$type[1]`* has the most expensive item(s) at $`r answer2$cost[1]`.  Those item(s) are `r answer2$itemName`.

3. At the Brodhead Center, how many of the entrees (found in the `menuType` variable) cost eight dollars?

```{r}
brodhead %>% 
  filter(cost == 8, menuType == "entree") %>% 
  select(name, menuType, itemName, cost)
```


4. The head of Duke dining is considering reducing prices at the Brodhead Center. Using what we've learned in class, write a line of code that will calculate a new variable (`halfPrice`) that contains items at half price.

```{r}
brodhead %>% 
  mutate(halfPrice = cost / 2) %>% 
  select(name, itemName, cost, halfPrice)
```


5. How many entrees are in the dataset (`menuType` variable)? How many desserts 
```{r}
brodhead %>% 
  count(menuType)
```
